Curve counting in genus one: elliptic singularities and relative geometry

### 2019-06-26 Reduced Gromov-Witten Theory of Hypersurfaces in Genus One.tex
minor changes to 5.3.3.

The next step is to construct $\widetilde\Ecal$ as a logarithmic modification of $\Ecal$, and to do this we need to understand the tropicalisation of the map $\widetilde\Dcal \to \Ecal$. In fact, it is more convenient to consider the \emph{extended tropicalisation} \cite{}. We begin by fixing notation. Associated to the combinatorial type $\Delta$, we have stable vertices $\sqC_0,\sqC_1,\ldots,\sqC_r$ with $\sqC_0$ corresponding to the core and $\sqC_1,\ldots,\sqC_r$ corresponding to stable vertices which lie on the raidus (see Lemma \ref{} above). For $i \in \{1,\ldots,r\}$ let $e_i$ denote the sum of the edge lengths separating $\sqC_i$ from $\sqC_0$. These form co-ordinates on the cone $\sigma$, and in these co-ordinates $\tau$ is the ray given by the equations $e_1=\ldots=e_r$. Moreover, the $e_i$ define functions on any cone in the star of $\sigma$.

Now consider the extended tropicalisation of $\widetilde\Dcal$. This is obtained inside the extended tropicalisation of $\widetilde\VZ_{1,\alpha}(\PP^N|H,d)$ by intersecting the star of $\tau$ with the locus $e_1=\ldots=e_r=\infty$. On the other hand, the extended tropicalisation of $\Ecal$ is obtained inside the extended tropicalisation of $\ol\Mcal_{1,\alpha}(\PP^N|H,d)$ by intersecting the star of $\sigma$ with the same locus. It is therefore clear that the extended tropicalisation of $\widetilde\Dcal$ is a subdivision of that of $\Ecal$. Moreover we claim that we can describe this subdivision explicitly.

Consider any cone $\rho$ in the star of $\tau$. This is defined by a number of equalities and inequalities involving edge lengths coming from the combinatorial type on $\rho$. Amongst these edge lengths are $e_1,\ldots,e_r$ which corresponding to the separating nodes which persist on all of $\widetilde\Dcal$; there are also other edge lengths, corresponding to nodes over $\rho$ which are smoothed out on $\tau$. We first deal with the (in)equalities relating distances of vertices obtained by degenerating $\sqC_1,\ldots,\sqC_r$. By the definition of a well-spaced tropical map, we see that these are all of the form
\begin{equation*} f + e_i = g + e_j \end{equation*}
where $f$ and $g$ are positive linear combinations of additional edge length parameters which are set to zero in $\tau$. We may now pass to the subcone of $\rho$ obtained by imposing the additional equations $e_1=\ldots=e_r$. Note that this is \emph{not} a subcone of $\tau$ in general, since there are additional edge lengths which must be set to zero if one wishes to recover $\tau$. The additional equations mean that we can rewrite our previous (in)equalities in the form
\begin{equation*} f = g \end{equation*}
and this is now well-defined in the limit $e_1=\ldots=e_r=\infty$. We thus obtain a cone in the subdivision of the extended tropicalisation of $\Ecal$, and all such cones are obtained in this manner.

@@

It remains to deal with the (in)equalities relating distances of vertices obtained by degenerating $\sqC_0$. These are not affected at all in the limit $e_1=\ldots=e_r=\infty$ and so the effect is to replace the factor of $\Ecal$ attached to the vertex $\sqC_0$
\begin{equation*} \ol\Mcal^{\leftrightarrow}_{\Gamma_{\tiny{\sqC_0}}}(P|H_0+H_\infty) \end{equation*}
with the radially aligned rubber space:
\begin{equation*} \VZ^{\leftrightarrow}_{\Gamma_{\tiny{\sqC_0}}}(P|H_0+H_\infty).\end{equation*}
The factorisation condition here is accounting for the ``fibre-direction'' factorisation; the ``base-direction'' factorisation has not entered the picture yet, and will appear when we come to describe $\Dcal$.

>>>>

The next step is to construct $\widetilde\Ecal$ as a logarithmic modification of $\Ecal$, and to do this we need to describe the tropicalisation of the map $\widetilde\Dcal \to \Ecal$. In fact, it is more convenient to consider the \emph{extended tropicalisation} \cite{}. We begin by fixing notation. Associated to the combinatorial type $\Delta$, we have stable vertices $\sqC_0,\sqC_1,\ldots,\sqC_r$ with $\sqC_0$ corresponding to the core and $\sqC_1,\ldots,\sqC_r$ corresponding to stable vertices which lie on the raidus (see Lemma \ref{} above). For $i \in \{1,\ldots,r\}$ let $e_i$ denote the sum of the edge lengths separating $\sqC_i$ from $\sqC_0$. These give functions on the cone $\sigma$, and $\tau$ is the ray given by the equations $e_1=\ldots=e_r$. Note that the $e_i$ also define functions on any cone in the star of $\sigma$.

Now consider the extended tropicalisation of $\widetilde\Dcal$. This is obtained inside the extended tropicalisation of $\widetilde\VZ_{1,\alpha}(\PP^N|H,d)$ by intersecting the star of $\tau$ with the locus $e_1=\cdots=e_r=\infty$. On the other hand, the extended tropicalisation of $\Ecal$ is obtained inside the extended tropicalisation of $\ol\Mcal_{1,\alpha}(\PP^N|H,d)$ by intersecting the star of $\sigma$ with the same locus. It is therefore clear that the extended tropicalisation of $\widetilde\Dcal$ is a subdivision of that of $\Ecal$. Moreover we claim that we can describe this subdivision explicitly.

Consider any cone $\rho$ in the star of $\tau$. Since the tropicalisation of $\widetilde\VZ_{1,\alpha}(\PP^N|H,d)$ is a subdivision of the tropicalisation of $\ol\Mcal_{1,\alpha}(\PP^N|H,d)$ there is a unique minimal cone in the star of $\sigma$ which contains $\rho$. We denote this cone by $\theta$, and remark that $\theta$ arises from a choice of combinatorial type which specialises the combinatorial type $\Delta$ defining $\sigma$. On $\theta$ we have the edge length functions $e_1,\ldots,e_r$ as well as additional edge lengths corresponding to nodes on $\theta$ which are smoothed out on $\sigma$.

Now, $\rho$ is defined inside $\theta$ by a number of equalities and inequalities involving these edge lengths. These (in)equalities relate the distances to the core of various vertices of the dual graph. There are three cases we must consider:
\begin{enumerate}
\item both vertices are components of $\sqC_1 \cup \ldots \cup \sqC_r$;
\item both vertices are components of $\sqC_0$;
\item one vertex is a component of $\sqC_0$ and one is a component of $\sqC_1 \cup \ldots \cup\sqC_r$.
\end{enumerate}
We start with case (1). By the definition of a well-spaced tropical map, we see that these are all of the form
\begin{equation*} f_i + e_i = f_j + e_j \end{equation*}
where $f_i$ and $f_j$ are positive linear combinations of additional edge length parameters which are set to zero in $\tau$. We may now pass to the subcone of $\rho$ obtained by imposing the additional equations $e_1=\ldots=e_r$. Note that this is \emph{not} a subcone of $\tau$ in general, since there are additional edge lengths which must be set to zero if one wishes to recover $\tau$. The additional equations mean that we can rewrite our previous (in)equalities in the form
\begin{equation*} f_i = f_j \end{equation*}
and these define a cone in the limiting locus $e_1=\ldots=e_r=\infty$. By definition this is one of the cones in the subdivision of the extended tropicalisation of $\Ecal$, and all such cones are obtained in this manner.

@@

We now deal with Case (2). These (in)equalities are not affected at all in the limit $e_1=\ldots=e_r=\infty$ and so the effect is to replace the factor of $\Ecal$ attached to the vertex $\sqC_0$
\begin{equation*} \ol\Mcal^{\leftrightarrow}_{\Gamma_{\tiny{\sqC_0}}}(P|H_0+H_\infty) \end{equation*}
with the radially aligned rubber space:
\begin{equation*} \VZ^{\leftrightarrow}_{\Gamma_{\tiny{\sqC_0}}}(P|H_0+H_\infty).\end{equation*}
The factorisation condition here is accounting for the ``fibre-direction'' factorisation; the ``base-direction'' factorisation has not entered the picture yet, and will appear when we come to describe $\Dcal$.

It remains to discuss Case (3). Here we have a relation of the form $f_i = f_j + e_j$, so in the limit $e_j =\infty$ there is no limiting cone. Therefore these relations to not appear in the modification of $\Ecal$.
